# Reproducible Research: Peer Assessment 1


## Loading and preprocesing data

### Loading

```{r}
packages <- c("data.table", "ggplot2", "xtable", "VIM")
sapply(packages, require, character.only=TRUE, quietly=TRUE)
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, destfile = "Dataset.zip", method = "curl")
unzip("Dataset.zip")
activity <- read.csv("./activity.csv")
```

### Process  the data

```{r}
activity$date <- as.Date(activity$date)
length(activity$steps)
summary(activity)

```


## What is mean total number of steps taken per day?

> For this part ignore the missing values in
> the dataset.
>
> 1. Make a histogram of the total number of steps taken each day
>
> 2. Calculate and report the **mean** and **median** total number of steps taken per day

Aggregate the number of steps taken each day.
Days with missing values (`NA`) will have `NA` when aggregated.

```{r}
m<-tapply(activity$steps, activity$date,sum)
edit(head(m))

```

Plot a histogram of the total number of steps taken each day. (for understand variability between days)

```{r }
hist(m)
```

Calculate the mean and median total number of steps taken per day **before imputing**.

```{r, results='asis'}
summary(m)
```



## What is the average daily activity pattern?

> 1. Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
>
> 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

Aggregate the average number of steps taken by 5-minute interval.

```{r}
average_steps<-data.frame(cbind(activity$interval,tapply(activity$steps, activity$interval, mean, na.rm = TRUE)))
colnames(average_steps) <- c("interval","steps")

```

Plot a time series of the 5-minute interval and the average number of steps taken across all days.

```{r timeseriesStepsTakenEachInterval}
q<-ggplot(data=average_steps,aes(x=interval,y=steps)) +
  geom_line()
print(q)
```


## Imputing missing values

1. Calculated missing

```{r}
table(is.na(activity$steps))
```

2. Filling in all dataset

Use the mean of interval, to replace the missing values.

```{r}
mfixed<-activity
mfixed[is.na(mfixed[, 1]), 1]<-average_steps[is.na(mfixed[, 1]),2]
```

Create a histogram of the new dataset, which has the NAs replaced with the 5-minute means.

```{r}
qplot(date, weight=mfixed$steps, data=mfixed, geom="histogram")
```

Calculate and report the mean and media values. We can see that the values differ from the initial calculation, and have higher value. It is also observed in the graph above. Most days have an increased number of mean of steps, since the NAs have been replaced by the 5-minute interval mean.

```{r}
mean(tapply(mfixed$steps, mfixed$date, sum, na.rm = TRUE))
median(tapply(mfixed$steps, mfixed$date, sum, na.rm = TRUE))
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
library(lattice)
mfixed$day<-as.factor(ifelse(weekdays(mfixed$date) %in% c("Saturday","Sunday"),"Weekend","Weekday"))
```
Plot the 5-minute average of steps, by weekday/weekend.
```{r}
xyplot(steps ~ interval | day, aggregate(steps ~ interval + day, mfixed, FUN = mean), 
layout = c(1, 2), type = "l", group=day)
```


